Distance Sampling in \proglang{R}

### 2015-12-05 paper.Rmd
more clarification of $K$/$T$

Having obtained the abundance in the covered area, we can then scale-up to the study area:
$$
\hat{N} = \frac{A}{a} \hat{N}_\text{C},
$$
where $A$ is the area of the study region to extrapolate the abundance estimate to and $a$ is the covered area. For line transects $a=2wL$ (twice the truncation distance multiplied by the total length of transects surveyed) and for points $a=\pi w^2 K$.

*have L and K been previously defined?*

@@

Sampling uncertainty can be characterised (rather than as simply $n$) by the *encounter rate*: the number of objects per unit transect. When covariates are not included in the detection function we can define the encounter rate as $n/L$ for line transects (where $L$ is the total line length) or $n/T$ for point transects (where $T$ is the total number of visits to all points) *earlier eqn uses notation $K$ for number of points*. When covariates are included in the detection function, it is recommended that we substitute the $n$ in the previous expressions with the estimated abundance $\hat{N}$ as this will take into account the effects of the covariates.

For line transects, by default, \pkg{Distance} uses a variation of the estimator "R2" from @Fewster:2009ku replacing number of observations per sample with the estimated abundance per sample [@Innes:2002ka; @Marques:2003vb]:
$$
\widehat{\text{Var}}_{\text{encounter},R2} = \frac{K}{L^2(K-1)} \sum_{k=1}^{K} l_k^2 \left( \frac{N_k}{l_k} - \frac{N}{L}\right)^2,
$$
where $l_k$ are the lengths of the $K$ transects (such that $L = \sum_{k=1}^K l_k$). Whereas for points we use estimator "P3" from @Fewster:2009ku but again replacing $n$ by $\hat{N}$ to obtain the following estimator:
$$
\widehat{\text{Var}}_{\text{encounter},P3} = \frac{1}{T(K-1)} \sum_{k=1}^{K} t_k \left( \frac{N_k}{t_k} - \frac{N}{T}\right)^2,
$$
where $t_k$ is the number of visits to point $k$ and $T = \sum_{k=1}^K t_k$. *this messes with the $K$ notation even further*

>>>>

Having obtained the abundance in the covered area, we can then scale-up to the study area:
$$
\hat{N} = \frac{A}{a} \hat{N}_\text{C},
$$
where $A$ is the area of the study region to extrapolate the abundance estimate to and $a$ is the covered area. For line transects $a=2wL$ (twice the truncation distance multiplied by the total length of transects surveyed, $L$) and for points $a=\pi w^2 T$ (where $\pi w^2$ is the area of a single surveyed circle and $T$ is the total number of visits to the sampled points).

@@

Sampling uncertainty can be characterised (rather than as simply $n$) by the *encounter rate*: the number of objects per unit transect. When covariates are not included in the detection function we can define the encounter rate as $n/L$ for line transects (where $L$ is the total line length) or $n/T$ for point transects (where $T$ is the total number of visits to all points). When covariates are included in the detection function, it is recommended that we substitute the $n$ in the previous expressions with the estimated abundance $\hat{N}$ as this will take into account the effects of the covariates.

For line transects, by default, \pkg{Distance} uses a variation of the estimator "R2" from @Fewster:2009ku replacing number of observations per sample with the estimated abundance per sample [@Innes:2002ka; @Marques:2003vb]:
$$
\widehat{\text{Var}}_{\text{encounter},R2} = \frac{K}{L^2(K-1)} \sum_{k=1}^{K} l_k^2 \left( \frac{N_k}{l_k} - \frac{N}{L}\right)^2,
$$
where $l_k$ are the lengths of the $K$ transects (such that $L = \sum_{k=1}^K l_k$). Whereas for points we use estimator "P3" from @Fewster:2009ku but again replacing $n$ by $\hat{N}$ to obtain the following estimator:
$$
\widehat{\text{Var}}_{\text{encounter},P3} = \frac{1}{T(K-1)} \sum_{k=1}^{K} t_k \left( \frac{N_k}{t_k} - \frac{N}{T}\right)^2,
$$
where $t_k$ is the number of visits to point $k$ and $T = \sum_{k=1}^K t_k$ (the total number of visits to all points is the sum of the visits to each point).
